---
title: "User's Guide for A Data Simulation Function For Any Multiple Linear model"
author: "Chenguang Pan"
date: "2023-04-21"
output:
  pdf_document:
    toc: false
    toc_depth: 4
    number_sections: true
    keep_tex: true
    highlight: tango
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = TRUE,
  prompt = TRUE,
  comment = '',
  collapse = TRUE,
  warning = FALSE,
  message = FALSE)
```   

## Project Address  

I synced the progress of this function on the Github. You can check it here if interested: https://github.com/cgpan/simdata  


## Data simulation function   

This function is to generate data from any given **multiple linear regression model**. Before running this function, please make sure you installed the package `mvtnorm`. Please run the latest version  `03_Data_simulation_V1.2.R`.  

There are several arguments a user need to know:  

  - `size=500`, the number of generated observations is set to 500 by default. You can give any number you want;  
  - `beta_=`, you must give a coefficient array. It should be in the sequence of c($\beta_0$,$\beta_1$,$\beta_2$,$\beta_3$...), which means the interception should be first numeric item;  
  - `pred_means=`, since this function mainly generate data from a multivariate norm distribution, you can give the variables means or it will use 0 by default;  
  - `pre_cov=`, you can give the variance-covariance matrix, or it will use the variance =1, covariance =0 matrix by default;
  - `error_df=1`, the error will be generated from a normal distribution with mean=0 and sd=1, you can set the sd of error based on your theory;  
  
The function will return a `dataframe` not a matrix, with the columns of `Y`, `X1`, `X2`,...
  
## The source code of this function  
```{r}
# before running the function , please load the package first
library(mvtnorm)
# write a function to generate data
dat_gen <- function(size=500, # datasize sets to 500 by default,you can change to any size
                    beta_ = NULL, # necessary, input the betas in the seuqnce of c(b0, b1, b2...)
                    pred_means = NULL, # not necessary,input the means of each variables
                    pred_cov = NULL, # not necessary, input the cov matrix of variables
                    error_sd = 1){
  if (is.null(beta_) == TRUE){
    # if user did not give a coefficients array, return warning.
    return("Error: You need to give the coefficients in the sequence of beta0, beta1,...")
  }else {
    if (is.null(pred_means)== TRUE){
      # if user did not give means of each variables,
      # use 0 as means by default
      predictor_nums <- length(beta_)
      if (is.null(pred_cov)== TRUE){
        print("Using the variance 1 and covariance 0 by default")
        X <- rmvnorm(n=size,sigma=diag(predictor_nums-1))
        Error <- as.matrix(rnorm(n=size, mean = 0, sd=error_sd))
        X_aug <- cbind(rep(1,nrow(X)), X)
        Y <- X_aug %*% as.matrix(beta_) + Error
        out_data <- cbind(Y,X)
      } else{ ### user gives the predictors covariance matrix
        X <- rmvnorm(n=size,sigma=pred_cov)
        Error <- as.matrix(rnorm(n=size, mean = 0, sd=error_sd))
        X_aug <- cbind(rep(1,nrow(X)), X)
        Y <- X_aug %*% as.matrix(beta_) + Error
        out_data <- cbind(Y,X)
      }
    }else{ ## user gives the means of predictors 
      if (is.null(pred_cov)==TRUE){
        print("Using the variance 1 and covariance 0 by default")
        X <- rmvnorm(n=size,mean= pred_means,sigma=diag(predictor_nums-1))
        Error <- as.matrix(rnorm(n=size, mean = 0, sd=error_sd))
        X_aug <- cbind(rep(1,nrow(X)), X)
        Y <- X_aug %*% as.matrix(beta_) + Error
        out_data <- cbind(Y,X)
      } else{ ### user gives the predictors covariance matrix
        X <- rmvnorm(n=size,mean= pred_means,sigma=pred_cov)
        Error <- as.matrix(rnorm(n=size, mean = 0, sd=error_sd))
        X_aug <- cbind(rep(1,nrow(X)), X)
        Y <- X_aug %*% as.matrix(beta_) + Error
        out_data <- cbind(Y,X)
      }
    }
    # give columns names in "Y","X1","X2",...
    n_ = predictor_nums - 1
    x_vars <- c("Y")
    for (i in 1:n_) {
      x_vars[i+1] <- paste0("X",i)
    }
    colnames(out_data) <- x_vars
    return(as.data.frame(out_data))
  }
}
```  

## Test of this function    
### Step 1. Run a linear model on real data  

Let's use the R-built-in dataset `mtcars` to test this function. First, I run a linear model as follows:
```{r}
# test
data(mtcars)
# Fit a linear regression model to predict miles per gallon (mpg) based on horsepower (hp)
model <- lm(mpg ~ hp+wt, data = mtcars)

# View the summary of the model
summary(model)
```  
The coefficients are `37.227`, `-0.032`, and `-3.878`. The standard deviation of error term is `r summary(model)$sigma`. Let's plug this value in the function and generate a sample with size of 500.    

### Step 2. Data simulation using the function above  

Note, in this step, I did not give the variance- covariance matrix. The function will use the var=1 and cov=0 by defalt.  

```{r}
set.seed(666)
sim_data <- dat_gen(size=500,
                     beta_=c(37.23,-0.03177295,3.87783074),
                     error_sd = summary(model)$sigma)
head(sim_data)
```  
The data looks good. Now, run linear model on this simulated dataset.  

### Step 3. Run a linear model on the simulated data.  

```{r}
sim_model <- lm(Y ~ X1 +X2, data = sim_data)
summary(sim_model)
```
Note, the first predictor become non-significant. We might want to input the covariance matrix of the variables to make data simulation more accurate. Let's try it again.


```{r}
# get the var-cov matrix
data(mtcars)
cov_m <- cov(mtcars[,c("hp","wt")])
# input the cov-matrix into the function and re-run the steps above
set.seed(666)
sim_data <- dat_gen(size=500,
                    beta_=c(37.23,-0.03177295,3.87783074),
                    pred_cov = cov_m,
                    error_sd = summary(model)$sigma)
sim_model <- lm(Y ~ X1 +X2, data = sim_data)
summary(sim_model)
```  
Now comparing the result from the model on the real data. these simulated data looks good!  




